02 Data analysis

### 2024-12-07 manuscript/manuscript.qmd
update code for latest tidygam which introduced soft-breaking changes

# Get model predictions for 'proficiency', assign them to 'gam_2_preds'
gam_2_preds <- predict_gam(
  gam_2, length_out = 25,
  series = "proficiency",
  exclude_terms = c("s(subj)")
)
```

@@

```{r}
#| label: gam-2-preds-t
# Get model predictions for 'time', assign them to 'gam_2_preds_t'
gam_2_preds_t <- predict_gam(
  gam_2, length_out = 25,
  series = "time",
  exclude_terms = c("s(subj)")
)
```

@@

```{r}
#| label: gam-3-preds
# Get model predictions, assign them to object 'gam_3_preds'
gam_3_preds <- predict_gam(gam_3, exclude_terms = "s(timep,subj)")
gam_3_preds
```

@@

```{r}
#| label: gam-3-preds-2
# Get model predictions, separate group_vowel column
gam_3_preds_2 <- predict_gam(
  gam_3,
  length_out = 25,
  exclude_terms = "s(timep,subj)",
  separate = list(group_vowel = c("group", "vowel"))
)
gam_3_preds_2
```

>>>>

# Get model predictions for 'proficiency', assign them to 'gam_2_preds'
gam_2_preds <- predict_gam(
  gam_2, length_out = 25,
  series = "proficiency",
  exclude_terms = c("s(subj)"),
  # Need to pick a single value for excluded terms due to limitations of mgcv
  # Any value will work. Since the term is excluded, picking a specific value
  # has no effect on the predictions.
  values = c(subj = "s1")
)
```

@@

```{r}
#| label: gam-2-preds-t
# Get model predictions for 'time', assign them to 'gam_2_preds_t'
gam_2_preds_t <- predict_gam(
  gam_2, length_out = 25,
  series = "time",
  exclude_terms = c("s(subj)"),
  # Need to pick a single value for excluded terms due to limitations of mgcv
  # Any value will work. Since the term is excluded, picking a specific value
  # has no effect on the predictions.
  values = c(subj = "s1")
)
```

@@

```{r}
#| label: gam-3-preds
# Get model predictions, assign them to object 'gam_3_preds'
gam_3_preds <- predict_gam(gam_3, exclude_terms = "s(timep,subj)", values = c(subj = "es_3"))
gam_3_preds
```

@@

```{r}
#| label: gam-3-preds-2
# Get model predictions, separate group_vowel column
gam_3_preds_2 <- predict_gam(
  gam_3,
  length_out = 25,
  exclude_terms = "s(timep,subj)",
  values = c(subj = "es_1"),
  separate = list(group_vowel = c("group", "vowel"))
)
gam_3_preds_2
```

### 2024-12-07 manuscript/manuscript.tex
update code for latest tidygam which introduced soft-breaking changes

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

\includegraphics{manuscript_files/figure-pdf/fig-dat1-1.pdf}

@@

\includegraphics{manuscript_files/figure-pdf/fig-gam-1-1.pdf}

@@

\includegraphics{manuscript_files/figure-pdf/fig-d1-dag-1.pdf}

@@

\includegraphics{manuscript_files/figure-pdf/fig-d1-dag-2.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Get model predictions for \textquotesingle{}proficiency\textquotesingle{}, assign them to \textquotesingle{}gam\_2\_preds\textquotesingle{}}
\NormalTok{gam\_2\_preds }\OtherTok{\textless{}{-}} \FunctionTok{predict\_gam}\NormalTok{(}
\NormalTok{  gam\_2, }\AttributeTok{length\_out =} \DecValTok{25}\NormalTok{,}
  \AttributeTok{series =} \StringTok{"proficiency"}\NormalTok{,}
  \AttributeTok{exclude\_terms =} \FunctionTok{c}\NormalTok{(}\StringTok{"s(subj)"}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{manuscript_files/figure-pdf/fig-gam-2-preds-1.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Get model predictions for \textquotesingle{}time\textquotesingle{}, assign them to \textquotesingle{}gam\_2\_preds\_t\textquotesingle{}}
\NormalTok{gam\_2\_preds\_t }\OtherTok{\textless{}{-}} \FunctionTok{predict\_gam}\NormalTok{(}
\NormalTok{  gam\_2, }\AttributeTok{length\_out =} \DecValTok{25}\NormalTok{,}
  \AttributeTok{series =} \StringTok{"time"}\NormalTok{,}
  \AttributeTok{exclude\_terms =} \FunctionTok{c}\NormalTok{(}\StringTok{"s(subj)"}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{manuscript_files/figure-pdf/fig-gam-2-preds-t-1.pdf}

@@

\includegraphics{manuscript_files/figure-pdf/fig-dat2-1.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Get model predictions, assign them to object \textquotesingle{}gam\_3\_preds\textquotesingle{}}
\NormalTok{gam\_3\_preds }\OtherTok{\textless{}{-}} \FunctionTok{predict\_gam}\NormalTok{(gam\_3, }\AttributeTok{exclude\_terms =} \StringTok{"s(timep,subj)"}\NormalTok{)}
\NormalTok{gam\_3\_preds}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 99 x 6
   group_vowel timep      eud      se  lower_ci upper_ci
   <fct>       <dbl>    <dbl>   <dbl>     <dbl>    <dbl>
 1 es.i          1    2.32    0.0155   2.29       2.35  
 2 es.u          1    2.02    0.0155   1.99       2.05  
 3 es.a          1    1.53    0.00809  1.52       1.55  
 4 es_0.i        1    0.0930  0.0473   0.000345   0.186 
 5 es_0.u        1    0.154   0.0476   0.0611     0.248 
 6 es_0.a        1   -0.00855 0.0410  -0.0888     0.0717
 7 es_1.i        1    1.10    0.0470   1.00       1.19  
 8 es_1.u        1    1.45    0.0447   1.36       1.53  
 9 es_1.a        1    1.28    0.0410   1.20       1.36  
10 es.i          1.8  2.32    0.0139   2.30       2.35  
# i 89 more rows
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Get model predictions, separate group\_vowel column}
\NormalTok{gam\_3\_preds\_2 }\OtherTok{\textless{}{-}} \FunctionTok{predict\_gam}\NormalTok{(}
\NormalTok{  gam\_3,}
  \AttributeTok{length\_out =} \DecValTok{25}\NormalTok{,}
  \AttributeTok{exclude\_terms =} \StringTok{"s(timep,subj)"}\NormalTok{,}
  \AttributeTok{separate =} \FunctionTok{list}\NormalTok{(}\AttributeTok{group\_vowel =} \FunctionTok{c}\NormalTok{(}\StringTok{"group"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{))}
\NormalTok{)}
\NormalTok{gam\_3\_preds\_2}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 234 x 7
   group vowel timep      eud      se  lower_ci upper_ci
   <chr> <chr> <dbl>    <dbl>   <dbl>     <dbl>    <dbl>
 1 es    i      1     2.32    0.0155   2.29       2.35  
 2 es    u      1     2.02    0.0155   1.99       2.05  
 3 es    a      1     1.53    0.00809  1.52       1.55  
 4 es_0  i      1     0.0930  0.0473   0.000345   0.186 
 5 es_0  u      1     0.154   0.0476   0.0611     0.248 
 6 es_0  a      1    -0.00855 0.0410  -0.0888     0.0717
 7 es_1  i      1     1.10    0.0470   1.00       1.19  
 8 es_1  u      1     1.45    0.0447   1.36       1.53  
 9 es_1  a      1     1.28    0.0410   1.20       1.36  
10 es    i      1.32  2.32    0.0148   2.30       2.35  
# i 224 more rows
\end{verbatim}

@@

\includegraphics{manuscript_files/figure-pdf/fig-gam-3-1.pdf}

>>>>

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\usepackage{bookmark}

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-dat1-1.pdf}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-gam-1-1.pdf}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-d1-dag-1.pdf}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-d1-dag-2.pdf}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Get model predictions for \textquotesingle{}proficiency\textquotesingle{}, assign them to \textquotesingle{}gam\_2\_preds\textquotesingle{}}
\NormalTok{gam\_2\_preds }\OtherTok{\textless{}{-}} \FunctionTok{predict\_gam}\NormalTok{(}
\NormalTok{  gam\_2, }\AttributeTok{length\_out =} \DecValTok{25}\NormalTok{,}
  \AttributeTok{series =} \StringTok{"proficiency"}\NormalTok{,}
  \AttributeTok{exclude\_terms =} \FunctionTok{c}\NormalTok{(}\StringTok{"s(subj)"}\NormalTok{),}
  \CommentTok{\# Need to pick a single value for excluded terms due to limitations of mgcv}
  \CommentTok{\# Any value will work. Since the term is excluded, picking a specific value}
  \CommentTok{\# has no effect on the predictions.}
  \AttributeTok{values =} \FunctionTok{c}\NormalTok{(}\AttributeTok{subj =} \StringTok{"s1"}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-gam-2-preds-1.pdf}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Get model predictions for \textquotesingle{}time\textquotesingle{}, assign them to \textquotesingle{}gam\_2\_preds\_t\textquotesingle{}}
\NormalTok{gam\_2\_preds\_t }\OtherTok{\textless{}{-}} \FunctionTok{predict\_gam}\NormalTok{(}
\NormalTok{  gam\_2, }\AttributeTok{length\_out =} \DecValTok{25}\NormalTok{,}
  \AttributeTok{series =} \StringTok{"time"}\NormalTok{,}
  \AttributeTok{exclude\_terms =} \FunctionTok{c}\NormalTok{(}\StringTok{"s(subj)"}\NormalTok{),}
  \CommentTok{\# Need to pick a single value for excluded terms due to limitations of mgcv}
  \CommentTok{\# Any value will work. Since the term is excluded, picking a specific value}
  \CommentTok{\# has no effect on the predictions.}
  \AttributeTok{values =} \FunctionTok{c}\NormalTok{(}\AttributeTok{subj =} \StringTok{"s1"}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-gam-2-preds-t-1.pdf}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-dat2-1.pdf}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Get model predictions, assign them to object \textquotesingle{}gam\_3\_preds\textquotesingle{}}
\NormalTok{gam\_3\_preds }\OtherTok{\textless{}{-}} \FunctionTok{predict\_gam}\NormalTok{(gam\_3, }\AttributeTok{exclude\_terms =} \StringTok{"s(timep,subj)"}\NormalTok{, }\AttributeTok{values =} \FunctionTok{c}\NormalTok{(}\AttributeTok{subj =} \StringTok{"es\_3"}\NormalTok{))}
\NormalTok{gam\_3\_preds}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 99 x 7
   group_vowel timep subj       eud     se lower_ci[,1] upper_ci[,1]
   <fct>       <dbl> <fct>    <dbl>  <dbl>        <dbl>        <dbl>
 1 es.i          1   es_3   2.32    0.0246       2.28         2.37  
 2 es.u          1   es_3   2.02    0.0246       1.97         2.07  
 3 es.a          1   es_3   1.53    0.0246       1.48         1.58  
 4 es_0.i        1   es_3   0.0930  0.0274       0.0394       0.147 
 5 es_0.u        1   es_3   0.154   0.0275       0.100        0.208 
 6 es_0.a        1   es_3  -0.00855 0.0246      -0.0568       0.0397
 7 es_1.i        1   es_3   1.10    0.0272       1.04         1.15  
 8 es_1.u        1   es_3   1.45    0.0261       1.39         1.50  
 9 es_1.a        1   es_3   1.28    0.0246       1.23         1.33  
10 es.i          1.8 es_3   2.32    0.0241       2.28         2.37  
# i 89 more rows
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Get model predictions, separate group\_vowel column}
\NormalTok{gam\_3\_preds\_2 }\OtherTok{\textless{}{-}} \FunctionTok{predict\_gam}\NormalTok{(}
\NormalTok{  gam\_3,}
  \AttributeTok{length\_out =} \DecValTok{25}\NormalTok{,}
  \AttributeTok{exclude\_terms =} \StringTok{"s(timep,subj)"}\NormalTok{,}
  \AttributeTok{values =} \FunctionTok{c}\NormalTok{(}\AttributeTok{subj =} \StringTok{"es\_1"}\NormalTok{),}
  \AttributeTok{separate =} \FunctionTok{list}\NormalTok{(}\AttributeTok{group\_vowel =} \FunctionTok{c}\NormalTok{(}\StringTok{"group"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{))}
\NormalTok{)}
\NormalTok{gam\_3\_preds\_2}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 234 x 8
   group vowel timep subj       eud     se lower_ci[,1] upper_ci[,1]
   <chr> <chr> <dbl> <fct>    <dbl>  <dbl>        <dbl>        <dbl>
 1 es    i      1    es_1   2.32    0.0246       2.28         2.37  
 2 es    u      1    es_1   2.02    0.0246       1.97         2.07  
 3 es    a      1    es_1   1.53    0.0246       1.48         1.58  
 4 es_0  i      1    es_1   0.0930  0.0274       0.0394       0.147 
 5 es_0  u      1    es_1   0.154   0.0275       0.100        0.208 
 6 es_0  a      1    es_1  -0.00855 0.0246      -0.0568       0.0397
 7 es_1  i      1    es_1   1.10    0.0272       1.04         1.15  
 8 es_1  u      1    es_1   1.45    0.0261       1.39         1.50  
 9 es_1  a      1    es_1   1.28    0.0246       1.23         1.33  
10 es    i      1.32 es_1   2.32    0.0244       2.28         2.37  
# i 224 more rows
\end{verbatim}

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-gam-3-1.pdf}}

### 2024-12-07 scripts/02_data_analysis.qmd
update code for latest tidygam which introduced soft-breaking changes

```{r}
predict_gam(
  d2_gam,
  length_out = 25,
  separate = list(group_vowel_o = c("group", "vowel")),
  exclude_terms = c("s(timep,subj)", "s(timep,subj):vowel_oi", "s(timep,subj):vowel_ou"),
  values = list(vowel_o = "a")
) %>%
  plot(series = "timep", comparison = "vowel") +
  facet_grid(cols = vars(group))
```

>>>>

```{r}
predict_gam(
  d2_gam,
  length_out = 25,
  separate = list(group_vowel_o = c("group", "vowel")),
  exclude_terms = c("s(timep,subj)", "s(timep,subj):vowel_oi", "s(timep,subj):vowel_ou"),
  values = list(vowel_o = "a", subj = "es_1")
) %>%
  plot(series = "timep", comparison = "vowel") +
  facet_grid(cols = vars(group))
```
